---
title: "Confounding"
subtitle: "Simulation of a Regression Analysis"
format: 
  html:
    toc: true
    brand: _brand.yml
editor: visual
---

## Data Generating Mechanism

Let's assume, we have a time invariant confounder $Depressiveness$ which affects two variables assessed at two time points $Absentism_{1}$, $Absentism_{2}$, $Screentime_{1}$, $Screentime_{2}$.

```{mermaid}
flowchart LR
  A1[Absentism₁] --> A2[Absentism₂]
  L1[Screentime₁] --> L2[Screentime₂]
  C[Depressiveness] --> A1
  C[Depressiveness] --> A2
  C[Depressiveness] --> L1
  C[Depressiveness] --> L2
```

> Note: We assume no effects of $Absentism_1$ on $Screentime_2$ and of $Screentime_1$ on $Absentism_2$.

## Simulate Data

```{r}
#| label: data sim
#| message: false
#| warning: false
library(tidyverse)
set.seed(2705)
n <- 1000

# Simulate confounder
data <- tibble(
  # Start with a purly independent variable
  Depressiveness = rnorm(n, mean = 0, sd = 1),
  # Add the two variables that are dependent on the confounder
  Absentism_1 = rnorm(n, mean = 0, sd = 1) + .25 * Depressiveness,
  Screentime_1 = rnorm(n, mean = 0, sd = 1) + .25 * Depressiveness,
  # Add the two variables that are dependent on the confounder and on each other 
  Absentism_2 = rnorm(n, mean = 0, sd = 1) + .25 * Depressiveness + .5*Absentism_1,
  Screentime_2 = rnorm(n, mean = 0, sd = 1) + .25 * Depressiveness + .5*Screentime_1,
)
```

## Regression Analysis

If we have data for the confounder simple multiple regression does the job of uncovering the confounding:

```{r}
library(sjPlot)
tab_model(
    lm(Absentism_2 ~ Screentime_1, data = data),
    lm(Absentism_2 ~ Depressiveness, data = data),
    lm(Absentism_2 ~ Screentime_1 + Depressiveness, data = data),
    show.ci = F,
    show.std = T
)

```

But what if we do not have data for the confounder? De-meaning/person-centering will help:

```{r}
data <- data %>%
  mutate(
    Absentism_1_pc = Absentism_1 - rowMeans(pick(Absentism_1, Absentism_2)),
    Absentism_2_pc = Absentism_2 - rowMeans(pick(Absentism_1, Absentism_2)),
    Screentime_1_pc = Screentime_1 - rowMeans(pick(Screentime_1, Screentime_2)),
    Screentime_2_pc = Screentime_2 - rowMeans(pick(Screentime_1, Screentime_2))
  )

tab_model(
    lm(Absentism_2_pc ~ Screentime_1_pc, data = data),
    show.ci = F,
    show.std = T
)
```

But how does this work? Maybe the following worked out example will help you to build an intuition:

![](img/DeMeaning.svg){.lightbox}

## Monte Carlo Simulation
### Simulate Data and Rerun Analyses
If you change the seed above and run the code again you will see that the results of the regression analysis change and even the *p*-values become significant sometimes. To test if the adjustment for `Depressiveness_₁` works *on avergage* a Monte Carlo Simulation can be used:

```{r}
#| label: monte carlo
library(broom)

# initiate two empty vector to store the results (p-value and regression weight)
b_depressiveness <- numeric(0)
p_depressiveness <- numeric(0)

# now run the Monte Carlo Simulation by repeating the data simulation and 
# regression analysis 10000 times (takes a minute) and store the results each time

for (i in 1:10000){
  # Simulate confounder
data_i <- 
  tibble(
    Depressiveness = rnorm(n, mean = 0, sd = 1),
    Absentism_1 = rnorm(n, mean = 0, sd = 1) + .25 * Depressiveness,
    Screentime_1 = rnorm(n, mean = 0, sd = 1) + .25 * Depressiveness,
    Absentism_2 = rnorm(n, mean = 0, sd = 1) + .25 * Depressiveness + .5*Absentism_1,
    Screentime_2 = rnorm(n, mean = 0, sd = 1) + .25 * Depressiveness + .5*Screentime_1)%>%
  mutate(
    Absentism_1_pc = Absentism_1 - rowMeans(pick(Absentism_1, Absentism_2)),
    Absentism_2_pc = Absentism_2 - rowMeans(pick(Absentism_1, Absentism_2)),
    Screentime_1_pc = Screentime_1 - rowMeans(pick(Screentime_1, Screentime_2)),
    Screentime_2_pc = Screentime_2 - rowMeans(pick(Screentime_1, Screentime_2))
  )

mod_i <- lm(Absentism_2_pc ~ Screentime_1_pc, data = data_i)

b_depressiveness[i] <- 
  tidy(mod_i) %>% 
    filter(term == "Screentime_1_pc") %>% 
    pull("estimate") %>% 
    as.numeric()

p_depressiveness[i] <- 
  tidy(mod_i) %>% 
    filter(term == "Screentime_1_pc") %>% 
    pull("p.value")
}
```

### Plot the results
```{r}
library(patchwork)
tibble(regression_weights = b_depressiveness) %>% 
  ggplot(aes(regression_weights)) +
    geom_histogram() +
    theme_minimal() +
    ggtitle("Regression Weights") +
tibble(p_values = p_depressiveness) %>% 
  ggplot(aes(p_values)) +
    geom_histogram() +
    theme_minimal() +
    ggtitle("p-values")
           
```

